---
title: "Week 4 Homework"
author: "Daniel Li"
date: "25/10/2021"
output:
  github_document:
    number_sections: false
    toc: true
    toc_depth: 3
---

# Task

Visualise time series data from last week's (week 3) data set on abundances

# Code

Load packages

```{r}
# Load packages
library(vroom)
library(tidyverse)
library(lubridate)
```

Read in the data

```{r}
##read in the data
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
wide_spp.1 <- vroom("../../data/Workshop 3/to_sort_pop_1.csv")
wide_spp.2 <- vroom("../../data/Workshop 3/to_sort_pop_2.csv")

## code to reshape data
## first join the data using full join - this will keep all of the columns
long_spp <- full_join(wide_spp.1, wide_spp.2) %>%
              ## pivot the joined data frame, using species, primary_threat, secondary_threat, tertiary_threat as ID columns
              ## and using names-pattern to pull out the population number
              ## and make a new column (called population) to store them in.
              ##Drop the NAs.
               pivot_longer(cols = -c(species,
                                      primary_threat,
                                      secondary_threat,
                                      tertiary_threat),
                            names_to = c("population", "date"),
                            names_pattern = "(.*)_(.*)",
                            values_drop_na = T,
                            values_to = "abundance")
```

Fix data frame

```{r}
# convert date column from chr to date
long_spp$date <- as_date(long_spp$date)
head(long_spp)

# remove NAs (commented out as NAs removed in full_join pivot earlier)
#long_spp <- drop_na(long_spp)
```

## Graph 1: Total abundance over time

```{r total-abundance-over-time}
# na.rm is superfluous as NAs already removed
abundance_year <- long_spp %>%
            group_by(date) %>%
            summarise("total.abundance" = sum(abundance, na.rm = T))
head(abundance_year)

abundance_year_gg <- ggplot(data = abundance_year, aes(x = date, y = total.abundance))

# check to ensure we dont have any issues in the data (like NAs)
which(is.na(abundance_year_gg$abundance))

abundance_year_gg + geom_line(aes(col = total.abundance)) +
                geom_point(aes(col = total.abundance)) +
                ggtitle("Total abundance over time") +
                theme(plot.title = element_text(hjust = 0.5)) +
                xlab("Date") +
                ylab("Total abundance") +
                labs(col = "Total abundance")
```

No need to export graph to file as R Markdown knits graphs into finished document

```{r, eval=FALSE}
# # export graph to file
# pdf("./Plots/homework/Total abundance over time.pdf", width = 6, height = 4)

#     abundance_year_gg + geom_line(aes(col = total.abundance)) +
#                         geom_point(aes(col = total.abundance)) +
#                         ggtitle("Total abundance over time") +
#                         theme(plot.title = element_text(hjust = 0.5)) +
#                         xlab("Date") +
#                         ylab("Total abundance") +
#                         labs(col = "Total abundance")

# dev.off()
```

## Graph 2: Population 1 vs 2 total abundance over time

Graphs generated for data overlayed and data on seperate graphs

```{r population-1-vs-2}
abundance_population_year <- long_spp %>%
            group_by(population, date) %>%
            summarise("total.abundance" = sum(abundance, na.rm = T))
head(abundance_population_year)

abundance_population_year_gg <- ggplot(data = abundance_population_year, aes(x = date, y = total.abundance))
abundance_population_year_gg + geom_line(aes(col = population)) +
                                geom_point(aes(col = population)) +
                                ggtitle("Total abundance by population") +
                                theme(plot.title = element_text(hjust = 0.5)) +
                                xlab("Date") +
                                ylab("Total abundance") +
                                labs(col = "Population")


abundance_population_year_gg + geom_line(aes(col = population)) +
                                geom_point(aes(col = population)) +
                                ggtitle("Total abundance by population") +
                                theme(plot.title = element_text(hjust = 0.5)) +
                                xlab("Date") +
                                ylab("Total abundance") +
                                labs(col = "Population") +
                                theme(legend.position = "none") +
                                facet_wrap(. ~ population)
```


## Graph 3: Bryum bornholmense abundance over time

```{r Bryum-bornholmense-abundance-over-time}
long_spp_bryum <- long_spp %>% filter(species == "Bryum bornholmense")

long_spp_bryum_gg <- ggplot(data = long_spp_bryum, aes(x = date, y = abundance))
long_spp_bryum_gg + geom_line(aes(col = abundance)) +
                        geom_point(aes(col = abundance)) +
                        ggtitle("Total abundance of Brynum bornholmense") +
                        theme(plot.title = element_text(hjust = 0.5)) +
                        xlab("Date") +
                        ylab("Abundance") +
                        labs(col = "Abundance") +
                        theme(legend.position = "none")
```